\protect Connecting every bit of knowledge: \\ The structure of Wikipedia's First Link Network

### 2016-04-27 writeup/2016-04wikipedia-network/wikipedia-network-revtex4-combined.tex
power law fit and descriptions complete

\usepackage{graphicx,epsfig,verbatim,enumerate}
\usepackage{amssymb,amsmath}
\usepackage{ifthen}

@@

The FLN's in-degree exhibits a decaying power law distribution where a few articles 
receive most direct first references, while most articles receive few or none.
The average in-degree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5.
Less than $1\%$ of articles have more than 100 direct first links and $75\%$ of articles
have fewer than 9. 
We fit the in-degree distribution against each article's rank with a log-log (base 10) linear regression model
(see Fig.~\ref{fig:degree distribution}).
We obtain a Pearson's Correlation Coefficient of $-0.98$
with a corresponding  power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$.

@@

Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{../graphics/traversals_per_article.png} 
  \caption{
    \textbf{Distribution of Traversal Visits.}
We fit the distribution of traversal visits against article rank with a log-log (base 10) linear regression model log-log model and uncover two regimes: 
with a cutoff rank of $10^{5}$.
The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
Coefficient of $-0.99$ 
with a corresponding power law rank exponent 
$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
The horizontal flattening around the highest
ranking articles is a result of the cyclic structure (see discussion on cycles).
The bottom regime has a Pearson's Correlation Coefficient of $-0.93$
with a corresponding power law rank exponent 
$\alpha = -0.579$ and a size exponent $\gamma = -0.727$.}
  \label{fig:Distribution of Visits}

@@

\begin{figure*}[tph!]
  \includegraphics[width=\textwidth]{../graphics/2_and_3_cycles.pdf}
  \caption{
    \textbf{Highest ranking 2-cycles and 3-cycles.}
We identify pairs of articles whose first links point to one another, forming
a 2-cycle. We then rank each pair of articles by the total number of 
traversal visits to gauge the most referenced groups of two articles linked
to each other. We find 2-cycles often capture synonyms or articles representing nearly the 
same concepts as opposed to distinct concepts. Similarly, we identify and 
rank 3-cycles to find they appear to capture three closely related ideas or synonyms.}
  \label{fig:cycles}
\end{figure*}
In Fig.~\ref{fig:Distribution of Visits}, we fit traversal visits against rank
to two regimes corresponding to power law rank exponents of 
$\alpha = -1.23 (\gamma = 0.187$) and $\alpha = -0.579 (\gamma = -0.727)$.
On the aggregate, the distribution suggests
a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none. The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{../graphics/funnels_distribution.png}
  \caption{
    \textbf{Distribution of Traversal Funnels.}
We fit the distribution with a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed number of traversal funnels. 
We find two regimes with the top regime (log(rank) < 4) 
well-explained by a linear fit, yielding Pearson's 
Correlation Coefficient of $-0.99$ and a 
corresponding power law rank exponent $\alpha = -1.08$
and a size exponent $\gamma = 0.074$. 
The bottom horizontal regime corresponds to the more than $99\%$ of articles
which hold zero traversal funnels.}
  \label{fig:Funnels Distribution}
\end{figure}

@@

Traversal funnels are comparatively resilient to changes in the FLN over time. While a few links can alter a cycle, potentially changing the number of traversal visits for an article, traversal funnels do not depend on cycles. Even when the particular links between articles change, the overall influence measured by traversal funnels is relatively resilient.

@@

\end{figure}
As a distribution, we find few articles influence the structure of the 
FLN. Only $17, 821$ articles have one or more traversal funnels, leaving
more than $99\%$ with none---most articles are recipients of 
the references flowing through the articles with at least one traversal funnel.
When fit to a log-log linear model we find the $99\%$ of articles with zero
traversal funnels form one regime (with $\log_{10}$(rank) less than 4).
The top regime, corresponding to the $17, 821$ articles with at least one 
traversal funnel strongly fits a linear model with
a Pearson's Correlation Coefficient of $-0.99$ and a 
corresponding power law rank exponent $\alpha = -1.08$
and a size exponent $\gamma = 0.074$ 
(see Fig.~\ref{fig:Funnels Distribution}). Even within the few articles
influencing the structure of the FLN, only a handful of these exert most of the 
control.

@@

\newpage

>>>>

\usepackage{graphicx,epsfig,verbatim,enumerate}
\usepackage{amssymb,amsmath}
\usepackage{ifthen}
\usepackage{hyperref}

% Fonts and typesetting settings
\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}
\linespread{1.05} % Palatino needs more space between lines
\usepackage{microtype}
\usepackage{textcomp}

@@

The FLN's in-degree exhibits a decaying power law distribution where a few articles 
receive most direct first references, while most articles receive few or none.
The average in-degree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5.
Less than $1\%$ of articles have more than 100 direct first links and $75\%$ of articles
have fewer than 9. 
We plot the in-degree distribution against each article's rank with a log-log (base 10) and fit the data to a power law 
distribution using the Maximum Likelihood Estimator with xmin as the indegree minimizing Kolmogorov-Smirnov distance.
(see Fig.~\ref{fig:degree distribution}).
We find xmin to be 5 links with a corresponding power law rank exponent $\alpha = 2.4$ and a size exponent $\gamma = 1.4$. 
Finally, we compare the fit against alternative distributions using a Kolmogorov-Smirnov Test. We find a significant log likelihood ratio favoring a  truncated power law ($R < 0, p=4.5\times10^{-13}$) given the maximum indegree is limited by the number of articles in the FLN. We find an inconclusive log likelihood ratio for a lognormal distribution and significant log likelihood ratios favoring a power law fit against an exponential ($R  > 0, p=0$) and against a stretched exponential ($R > 0, p = 3.0\times10^{-50}$).

@@

Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{../graphics/traversals_per_article.png} 
  \caption{
    \textbf{Distribution of Traversal Visits.}
    We plot the number of traversal visits for each article by rank on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
    distribution using the Maximum Likelihood Estimator with xmin as the number of traversal visits minimizing Kolmogorov-Smirnov distance. We find xmin to be 866 with a corresponding power law rank exponent $\alpha = 1.7$ and a size exponent $\gamma = 1.6$.
The horizontal flattening around the highest
ranking articles is a result of the cyclic structure (see discussion of Cycles ~\ref{cycles}).
}
  \label{fig:Distribution of Visits}
\end{figure}

@@

\begin{figure*}[tph!]
  \includegraphics[width=\textwidth]{../graphics/2_and_3_cycles.pdf}
  \caption{
    \textbf{Highest ranking 2-cycles and 3-cycles.}
We identify pairs of articles whose first links point to one another, forming
a 2-cycle. We then rank each pair of articles by the total number of 
traversal visits to gauge the most referenced groups of two articles linked
to each other. We find 2-cycles often capture synonyms or articles representing nearly the 
same concepts as opposed to distinct concepts. Similarly, we identify and 
rank 3-cycles to find they appear to capture three closely related ideas or synonyms.}
  \label{fig:cycles}
\end{figure*}
In Fig.~\ref{fig:Distribution of Visits}, we fit traversal visits against a power law distribution using the Maximum Likelihood Estimator with xmin as the number of traversal visits minimizing Kolmogorov-Smirnov distance. We find xmin to be 866 with a corresponding power law rank exponent $\alpha = 1.7$ and a size exponent $\gamma = 1.6$.
We also compared the fit against alternative distributions and found signficiant likelihood ratios favoring a power law fit to an exponential ($R>0, p=2.5\times10^{-60}$) and stretched exponential ($R > 0, p=.0001$).
Similarly to indegree, we found an inconclusive comparison against a lognormal distribution and a signficiant loglikelihood ratio favoring a truncated power law given the max number of traversal visits is restricted by the number of articles in the FLN. 
On the aggregate, the distribution of traversal visits suggests
a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none. The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{../graphics/funnels_distribution.png}
  \caption{
    \textbf{Distribution of Traversal Funnels.}
    We plot the number of traversal funnels for each article by rank on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
    distribution using the Maximum Likelihood Estimator with xmin as the number of traversal visits minimizing Kolmogorov-Smirnov distance. We find xmin to be 30 with a corresponding power law rank exponent $\alpha = 1.73$ and a size exponent $\gamma = 1.57$.
    The outlier on the top left is Philosophy with more traversal funnels than the second highest ranking article by two orders of magnitude.
}
  \label{fig:Funnels Distribution}
\end{figure}

@@

In contrast to traversal visits, traversal funnels are comparatively resilient to changes in the FLN over time. 
While a few links can alter a cycle, potentially displacing an article that was the recipient of references, traversal funnels remain stable even when cycles change.

@@

\end{figure}
As a distribution, we find few articles influence the structure of the 
FLN. Only $17, 821$ articles have one or more traversal funnels, leaving
more than $99\%$ with none---most articles are recipients of 
the references flowing through the articles with at least one traversal funnel.
When to a power law 
    distribution (using the Maximum Likelihood Estimator) with xmin as the number of traversal funnels minimizing Kolmogorov-Smirnov distance), we find xmin to be 30 with a corresponding power law rank exponent $\alpha = 1.73$ and a size exponent $\gamma = 1.57$
(see Fig.~\ref{fig:Funnels Distribution}). 
When compared against alternative distributions, we find inconclusive results for lognormal and truncated power law distributions and significant log likelihood ratios favoring a power law for exponential ($R>0, p=5.3\times10^{-6}$) and stretched exponential ($R>0, p = 1.8\times10^{-8}$) distributions.
Even within the few articles
influencing the structure of the FLN, only a handful of these exert most of the 
control.

@@

### 2016-04-27 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
power law fit and descriptions complete

@@

The FLN's in-degree exhibits a decaying power law distribution where a few articles 
receive most direct first references, while most articles receive few or none.
The average in-degree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5.
Less than $1\%$ of articles have more than 100 direct first links and $75\%$ of articles
have fewer than 9. 
We plot the in-degree distribution against each article's rank with a log-log (base 10) and fit the data to a power law 
distribution using the Maximum Likelihood Estimator with xmin as the indegree minimizing Kolmogorov-Smirnov distance.
(see Fig.~\ref{fig:degree distribution}).
We find xmin to be 5 links with a corresponding power law rank exponent $\alpha = 2.4$ and a size exponent $\gamma = 1.4$. 
Finally, we compare the fit against alternative distributions using a Kolmogorov-Smirnov Test. We find a significant log likelihood ratio favoring a  truncated power law ($R < 0, p=4.5\times10^{-13}$) given the limits on the maximum possible indegree.  We find an inconclusive log likelihood ratio for a lognormal distribution and significant log likelihood ratios favoring a power law fit to an exponential ($R  > 0, p=0$) and to a stretched exponential ($R > 0, p = 3.0\times10^{-50}$).

@@

Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{../graphics/traversals_per_article.png} 
  \caption{
    \textbf{Distribution of Traversal Visits.}
We fit the distribution of traversal visits against article rank with a log-log (base 10) linear regression model log-log model and uncover two regimes: 
with a cutoff rank of $10^{5}$.
The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
Coefficient of $-0.99$ 
with a corresponding power law rank exponent 
$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
The horizontal flattening around the highest
ranking articles is a result of the cyclic structure (see discussion on cycles).
The bottom regime has a Pearson's Correlation Coefficient of $-0.93$
with a corresponding power law rank exponent 
$\alpha = -0.579$ and a size exponent $\gamma = -0.727$.}
  \label{fig:Distribution of Visits}
\end{figure}

@@

\acknowledgments
MI is grateful to RJ for pointing out the Reddit post 
\cite{reddit}
describing how the majority of links lead to ``Philosophy,'' inspiring this research.

The authors are also grateful for the suggestions and input provided by Randall Harp.

PSD and CMD acknowledge support from NSF Big Data Grant $\#1447634$.

The authors also acknowledge support from the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.

\newpage

>>>>

%% :%s/{../graphics/{..graphics

@@

The FLN's in-degree exhibits a decaying power law distribution where a few articles 
receive most direct first references, while most articles receive few or none.
The average in-degree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5.
Less than $1\%$ of articles have more than 100 direct first links and $75\%$ of articles
have fewer than 9. 
We plot the in-degree distribution against each article's rank with a log-log (base 10) and fit the data to a power law 
distribution using the Maximum Likelihood Estimator with xmin as the indegree minimizing Kolmogorov-Smirnov distance.
(see Fig.~\ref{fig:degree distribution}).
We find xmin to be 5 links with a corresponding power law rank exponent $\alpha = 2.4$ and a size exponent $\gamma = 1.4$. 
Finally, we compare the fit against alternative distributions using a Kolmogorov-Smirnov Test. We find a significant log likelihood ratio favoring a  truncated power law ($R < 0, p=4.5\times10^{-13}$) given the maximum indegree is limited by the number of articles in the FLN. We find an inconclusive log likelihood ratio for a lognormal distribution and significant log likelihood ratios favoring a power law fit against an exponential ($R  > 0, p=0$) and against a stretched exponential ($R > 0, p = 3.0\times10^{-50}$).

@@

Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{../graphics/traversals_per_article.png} 
  \caption{
    \textbf{Distribution of Traversal Visits.}
    We plot the number of traversal visits for each article by rank on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
    distribution using the Maximum Likelihood Estimator with xmin as the number of traversal visits minimizing Kolmogorov-Smirnov distance. We find xmin to be 866 with a corresponding power law rank exponent $\alpha = 1.7$ and a size exponent $\gamma = 1.6$.
The horizontal flattening around the highest
ranking articles is a result of the cyclic structure (see discussion of Cycles ~\ref{cycles}).
}
  \label{fig:Distribution of Visits}
\end{figure}

@@

### 2016-04-27 writeup/wikipedia-network.tex
power law fit and descriptions complete

Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
  \caption{
    \textbf{Distribution of Traversal Visits.}
We fit the distribution of traversal visits against article rank with a log-log (base 10) linear regression model log-log model and uncover two regimes: 
with a cutoff rank of $10^{5}$.
The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
Coefficient of $-0.99$ 
with a corresponding power law rank exponent 
$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
The horizontal flattening around the highest
ranking articles is a result of the cyclic structure (see discussion on cycles).
The bottom regime has a Pearson's Correlation Coefficient of $-0.93$
with a corresponding power law rank exponent 
$\alpha = -0.579$ and a size exponent $\gamma = -0.727$.}
  \label{fig:Distribution of Visits}
\end{figure}

@@

\begin{figure*}[tph!]
  \includegraphics[width=\textwidth]{graphics/2_and_3_cycles.pdf}
  \caption{
    \textbf{Highest ranking 2-cycles and 3-cycles.}
We identify pairs of articles whose first links point to one another, forming
a 2-cycle. We then rank each pair of articles by the total number of 
traversal visits to gauge the most referenced groups of two articles linked
to each other. We find 2-cycles often capture synonyms or articles representing nearly the 
same concepts as opposed to distinct concepts. Similarly, we identify and 
rank 3-cycles to find they appear to capture three closely related ideas or synonyms.}
  \label{fig:cycles}
\end{figure*}
In Fig.~\ref{fig:Distribution of Visits}, we fit traversal visits against rank
to two regimes corresponding to power law rank exponents of 
$\alpha = -1.23 (\gamma = 0.187$) and $\alpha = -0.579 (\gamma = -0.727)$.
On the aggregate, the distribution suggests
a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none. The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
  \caption{
    \textbf{Distribution of Traversal Funnels.}
We fit the distribution with a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed number of traversal funnels. 
We find two regimes with the top regime (log(rank) < 4) 
well-explained by a linear fit, yielding Pearson's 
Correlation Coefficient of $-0.99$ and a 
corresponding power law rank exponent $\alpha = -1.08$
and a size exponent $\gamma = 0.074$. 
The bottom horizontal regime corresponds to the more than $99\%$ of articles
which hold zero traversal funnels.}
  \label{fig:Funnels Distribution}
\end{figure}

>>>>

Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
  \caption{
    \textbf{Distribution of Traversal Visits.}
    We plot the number of traversal visits for each article by rank on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
    distribution using the Maximum Likelihood Estimator with xmin as the number of traversal visits minimizing Kolmogorov-Smirnov distance. We find xmin to be 866 with a corresponding power law rank exponent $\alpha = 1.7$ and a size exponent $\gamma = 1.6$.
The horizontal flattening around the highest
ranking articles is a result of the cyclic structure (see discussion of Cycles ~\ref{cycles}).
}
  \label{fig:Distribution of Visits}
\end{figure}

@@

\begin{figure*}[tph!]
  \includegraphics[width=\textwidth]{graphics/2_and_3_cycles.pdf}
  \caption{
    \textbf{Highest ranking 2-cycles and 3-cycles.}
We identify pairs of articles whose first links point to one another, forming
a 2-cycle. We then rank each pair of articles by the total number of 
traversal visits to gauge the most referenced groups of two articles linked
to each other. We find 2-cycles often capture synonyms or articles representing nearly the 
same concepts as opposed to distinct concepts. Similarly, we identify and 
rank 3-cycles to find they appear to capture three closely related ideas or synonyms.}
  \label{fig:cycles}
\end{figure*}
In Fig.~\ref{fig:Distribution of Visits}, we fit traversal visits against a power law distribution using the Maximum Likelihood Estimator with xmin as the number of traversal visits minimizing Kolmogorov-Smirnov distance. We find xmin to be 866 with a corresponding power law rank exponent $\alpha = 1.7$ and a size exponent $\gamma = 1.6$.
We also compared the fit against alternative distributions and found signficiant likelihood ratios favoring a power law fit to an exponential ($R>0, p=2.5\times10^{-60}$) and stretched exponential ($R > 0, p=.0001$).
Similarly to indegree, we found an inconclusive comparison against a lognormal distribution and a signficiant loglikelihood ratio favoring a truncated power law given the max number of traversal visits is restricted by the number of articles in the FLN. 
On the aggregate, the distribution of traversal visits suggests
a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none. The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
  \caption{
    \textbf{Distribution of Traversal Funnels.}
    We plot the number of traversal funnels for each article by rank on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
    distribution using the Maximum Likelihood Estimator with xmin as the number of traversal visits minimizing Kolmogorov-Smirnov distance. We find xmin to be 30 with a corresponding power law rank exponent $\alpha = 1.73$ and a size exponent $\gamma = 1.57$.
    The outlier on the top left is Philosophy with more traversal funnels than the second highest ranking article by two orders of magnitude.
}
  \label{fig:Funnels Distribution}
\end{figure}
